---
title: "Initial Korean covid 19"
author: "Matthew Ramirez"
date: "5/30/2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```

#Introduction



### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female  
77 of the data entries were coded sex  as blank, so we omit these.  
state of a patient must be died or alive - we adjust that as well     
  
# Your Question - Is dying independent of sex?

 $H_0$: There is no difference in the amount of deaths per male and female.
 
 $H_A$: There is a difference between the bar and numerical charts that showcases the amount of deaths per male and female.

# Methods

Result is categorial variable, sex is categrical result  
Depends on sex, we use helper file  
cat~cat  

# Results

## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
  
This barchart shows that more Males died than females and more women lived than died.  
  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that the amount dead and alive for both males and females with the main result showing no difference in the amount of deaths.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for the chi-squared test is 0.001, so we reject the null hypothesis. 

The Fisher exact test shows that men are 2.3 times more likely to die thatn women in Korea.  

## Conclusions

Overall the barplot consists that men will have a higher death rate, numerical data shows how its an equal amount of death rates, and the inferential results show that the death rate for men is 2.3 times more likely than women.
